A full approximation storage multilevel method \\ for nonlinear variational inequalities

### 2023-07-07 paper/fascd.tex
First editing pass complete

\newcommand{\XX}{{\sffamily X}}

@@

To allow box constraints on all meshes in the hierarchy, let $\tilde{\mathcal{V}}^j$ denote the set of functions on the nodes of $\mathcal{T}^j$ with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.  Suppose that the finest-level \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$ satisfy
\begin{equation}
\underline{\gamma}^J \le \overline{\gamma}^J, \qquad \underline{\gamma}^J < +\infty, \qquad -\infty < \overline{\gamma}^J. \label{eq:fe:boxconstraintrequirements}
\end{equation}
\pef{Shouldn't there be some relation between the continuous bounds and the discrete ones?}
In analogy to the continuous data, also assume that $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ on those nodes of $\mathcal{T}^J$ which lie in $\partial_D \Omega$.  Define
\begin{equation}
\mathcal{K}^J = \left\{w\,:\,\underline{\gamma}^J \le w \le \overline{\gamma}^J \text{ on nodes of } \mathcal{T}^J, \, \text{ and } \, w|_{\partial_D\Omega} = g_D^J|_{\partial_D\Omega}\right\} \subset \mathcal{V}^J, \label{eq:fe:fineconstraintset}
\end{equation}
the finest-level \emph{constraint set} of admissible FE functions. This is a nonempty, closed, and convex subset of $\mathcal{V}^J$.  %As $\mathcal{V}^J$ functions never take values $\pm\infty$, inequality in \eqref{eq:fe:fineconstraintset} is strict where $\underline{\gamma}^J, \overline{\gamma}^J$ are infinite; there is no lower or upper constraint, respectively, at such nodes.

@@

While \eqref{eq:fe:vi} approximates \eqref{eq:boxdirichletvi}, we warn the reader that it may not technically be conforming for that VI.  That is, even if quadrature and other implementation details for $f^J$ are ignored, and despite the vector space inclusion $\mathcal{V}^J \subset \mathcal{V}$, we do not require or generally expect $\mathcal{K}^J \subset \mathcal{K}$.  Regardless of the smoothness of the continuum obstacles $\underline{\gamma}, \overline{\gamma}$, admissibility with respect to their FE interpolants $\underline{\gamma}^J, \overline{\gamma}^J$ will generally not imply continuum admissibility.\footnote{This caveat is especially relevant in geophysical situations where only nodal values of the obstacle are known.  Nodal elevation data generally misses the actual topographic detail in-between the nodes; compare \cite{Bueler2016}.}

@@

To define injection $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$, observe that the nodes of $\mathcal{T}^{j-1}$ are nodes of $\mathcal{T}^j$, and that for any values at $\mathcal{T}^j$ nodes there is a unique $\mathcal{V}^j$ function assuming those values.  For $z\in\mathcal{V}^j$ we define the injection $\iR z$ as the element of $\mathcal{V}^{j-1}$ with the same (coarse) nodal values.  Injection is monotonic: if $w \le z$ then $\iR w \le \iR z$.

\pef{I don't like the name ``monotone restrictions''. The word restriction is strongly associated with operating on the dual space. I would prefer to call them ``monotone injections'', since that is the matching operator.}
The monotone restrictions $\maxR,\minR$ \cite{GraeserKornhuber2009}, which will only be applied to nodal functions $\tilde{\mathcal{V}}^j$, are nonlinear maps which maximize and minimize, respectively, nodal values over the open supports of coarser-mesh basis functions (i.e.~over the \emph{stars} of the coarse vertices). These operators are constructed to enforce the property
\begin{equation}
\minR z \le z \le \maxR z,  \label{eq:fe:monotonerestrictionprops}
\end{equation}
which does not hold for the standard injection operator $\iR$.
To make the construction precise, let $x_q^j$ denote the nodes of mesh $\mathcal{T}^j$, with corresponding basis functions $\psi_q^j(x)$.  For a $\tilde\RR$-valued nodal function $z\in\tilde{\mathcal{V}}^j$ we define functions in $\tilde{\mathcal{V}}^{j-1}$:
\begin{align}
(\maxR z)(x_p^{j-1}) &= \max \{z(x_q^j) \,:\, \psi_p^{j-1}(x_q^j) > 0\}, \label{eq:fe:monotonerestrictions} \\
(\minR z)(x_p^{j-1}) &= \min \{z(x_q^j) \,:\, \psi_p^{j-1}(x_q^j) > 0\}. \notag
\end{align}
For standard refinement a coarse-mesh node $x_p^{j-1}$ is in the set of fine nodes $x_q^j$ such that $\psi_p^{j-1}(x_q^j)>0$, so \eqref{eq:fe:monotonerestrictionprops} follows.
On the other hand, $z\ge 0$ implies $\minR z \ge 0$ and $z \le 0$ implies $\maxR z \le 0$.  These ordering properties are used in the CD construction of the next two Sections.
\pef{It would be good to have a diagram here. I could write one in tikz, if you don't have one handy.}

@@

Let $\Omega = (0,L) \subset \RR^1$ and consider a two-level mesh hierarchy ($J=1$).  Let $\underline{\gamma}^1$ be a sawtooth function with minimum value zero, and suppose $\overline{\gamma}^1=C+\underline{\gamma}^1$ with $C>0$ which is strictly less than the maximum of $\underline{\gamma}^1$.  Figure \ref{fig:directRbad} shows such functions; here levels $\mathcal{T}^1$ and $\mathcal{T}^0$ have $5$ and $3$ nodes, respectively.  Ignoring Dirichlet boundary for simplicity, the constraint set $\mathcal{K}^1$ defined by \eqref{eq:fe:fineconstraintset} is nonempty.  However, various restriction schemes applied directly to $\underline{\gamma}^1,\overline{\gamma}^1$ are problematic.  Suppose we propose
    $$\underline{\gamma}^0, \overline{\gamma}^0 \equiv \iR \underline{\gamma}^1, \iR \overline{\gamma}^1. \qquad \text{\emph{[problematic]}}$$
Then $\underline{\gamma}^0=0$ and $\overline{\gamma}^0=C$ identically.  In that case the corresponding constraint set $\mathcal{K}^0$ is nonempty, but prolongation $Pw^0$ of an admissible solution of the coarse problem ($\underline{\gamma}^0 \le w^0 \le \overline{\gamma}^0$) is not admissible on the finer level: $Pw^0 \notin \mathcal{K}^1$.  On the other hand, suppose we propose
    $$\underline{\gamma}^0, \overline{\gamma}^0 \equiv \maxR \underline{\gamma}^1, \minR \overline{\gamma}^1. \qquad \text{\emph{[problematic]}}$$
The general motivation for this choice---compare \eqref{eq:fe:chilevels} below---is that $w^0 \ge \underline{\gamma}^0 \, \implies \, Pw^0 \ge \underline{\gamma}^1$ and $w^0 \le \overline{\gamma}^0 \, \implies \, Pw^0 \le \overline{\gamma}^1$.  However, while these facts are true, the difficulty with this choice is that the coarse-level admissible set $\mathcal{K}^0$ is now empty because the proposed lower bound $\underline{\gamma}^0$ actually exceeds the proposed upper bound $\overline{\gamma}^0$.
\end{example}

@@

\begin{example}  \label{ex:notfullcd}
Let $a > 0$ be an arbitrary positive number.  Consider a 2-level hierarchy ($J=1$), and suppose the lower defect constraint is the constant function $\underline{\chi}^1=-a$.  Note $\underline{\chi}^0=\maxR \underline{\chi}^1=-a$ also, so $\underline{\phi}^1=0$ and $\underline{\phi}^0=-a$.  Suppose $z^1$ is a ``sawtooth'' function which takes on value $-a$ on the coarse nodes and value $0$ on the other nodes.  Then, for any $\overline{\chi}^1\ge 0$, we have $\overline{\chi}^1 \ge 0 \ge z^1\ge \underline{\chi}^1$, so $z^j \in \mathcal{U}^1$.  However, $I_{1\to 0}^\ominus(z^1 - \underline{\chi}^1) = \minR(z^1 - \underline{\chi}^1) = 0$ identically.  From \eqref{eq:fe:unilateraldecompositionoperator}, $\Pi_1 z^1$ simplifies to $\Pi_1 z^1 = z^1 - \underline{\chi}^1$.  But then $\Pi_1 z^1$ attains a maximum $a\ge 0$.  This maximum can be chosen to exceed any upper defect obstacle $\overline{\phi}^1$, so $\Pi_1 z^1 \notin \mathcal{D}^1$.
\end{example}

@@

Because the problem is inequality-constrained, repeated application of Algorithm \ref{alg:fascd} to solve the finite-dimensional, box-constrained VI \eqref{eq:fe:vi} will not generally reduce the residual $f^J(w^J) - \ell^J$ to zero everywhere.  To construct appropriate convergence criteria for the discretization, note \eqref{eq:fe:vi} is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003}.  Let $x_p \in \mathcal{T}^J$ be a finest-level node, and $\psi_p(x)$ the corresponding basis function.  Let $\mathcal{X}^J=\big\{w\in\cV^J\,:\,w(x_p)=g_D^J(x_p) \text{ if } x_p\in \partial_D\Omega\big\}$, thus $\cK^J = \left\{\underline{\gamma}^J \le w \le \overline{\gamma}^J\right\} \cap \mathcal{X}^J$.  If $m_D$ is the number of nodes $x_p$ in the Dirichlet boundary, let $I_D$ be the $m_D\times m_J$ matrix which extracts Dirichlet nodal values; its transpose $I_D^\top$ injects Dirichlet boundary degrees of freedom into $(\mathcal{V}^J)'$.  Finally, write $a\perp b$ when $a,b \in \RR^n$ are \emph{complementary}, that is, when $a_i \ge 0$ and $b_i \ge 0$ and $a_i b_i = 0$ for $i=0,\dots,n-1$.  Then problem \eqref{eq:fe:vi}, which would be expressed VI$\left(f^J(\cdot)-\ell^J,\left[\,\underline{\gamma}^J,\overline{\gamma}^J\right] \cap \mathcal{X}^J\right)$ in the operations research literature \cite{FerrisPang1997}, is equivalent to the following MCP for the solution $u^J \in \mathcal{K}^J$, but with additional equality constraints:
\begin{align}
\underline{\gamma}^J \le u^J \le \overline{\gamma}^J \quad &\perp \quad f^J(u^J) - \ell^J - I_D^\top \lambda \label{eq:fe:mcp} \\
I_D u^J &= g_D^J\big|_{\partial_D\Omega} \notag
\end{align}
for some Lagrange multipliers $\lambda \in \RR^{m_D}$.
\pef{I wonder if this definition will confuse people, because they won't know how to plug in $\underline{\gamma}^J \le u^J \le \overline{\gamma}^J$ into your definition of $\perp$ ($\underline{\gamma}^J \le u^J \le \overline{\gamma}^J$ is not in $\RR^n$). Maybe it would be easier to just use the spelled-out version below, and not use this notation?}

@@

In summary, $\rNC : \mathcal{K}^J \to \RR^{m_J}$ is a computable nonlinear map defined using the discretization $f^J$ and the vectors $\ell^J$, $\underline{\gamma}^J$, and $\overline{\gamma}^J$.  A value $\rNC(w^J)_p$ is nonzero only at nodes where \eqref{eq:fe:mcp} is violated, either because $r(w^J)_p$ remains nonzero where the constraints are inactive, or because it has the wrong sign at an active node, or because the boundary condition is not satisfied.

@@

Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} were implemented in Python using the Firedrake library \cite{Rathgeberetal2016}, allowing $P_1$ or $Q_1$ elements and based on standard, global mesh refinement.  Here we show numerical results for the several example problems in Section \ref{sec:vi}, with a focus on the scaling of FASCD iterations as the mesh resolution increases.  Our essential observations are that each V-cycle dramatically reduces the residual norm, and that very few F-cycles solve most examples to within discretization error.

\begin{example} \label{ex:results:classical}
Consider the 2D classical, i.e.~unilateral and Laplacian, obstacle problem, namely Example \ref{ex:plaplacian} with $p=2$ and admissible set $\mathcal{K} = \{v \ge \psi\} \subset W^{1,2}(\Omega)$.  We consider two problems with very different coincidence (active) sets $\{x\in\Omega \,:\, u(x)=\psi(x)\}$, both over square $\Omega$.   The standard ``ball'' problem has an upper hemisphere obstacle $\psi$ \cite[Chapter 12]{Bueler2021}.  In the ``spiral'' problem from Example 7.1.1 in \cite{GraeserKornhuber2009}, $\psi$ is constructed to produce a more interesting coincidence set; see Figure \ref{fig:results:classical}.

For this experiment we used a coarsest mesh of $m_0=5^2$ nodes (32 triangles).  The finest mesh had $m_9=2049^2 \approx 4.2 \times 10^{6}$ nodes.  As our smoother we applied a single RS Newton iteration (Section \ref{sec:implementation}), with three (fixed) IC-preconditioned CG iterations to approximately solve the step equations.  The coarse solver was RS$+$LU, iterated to convergence.  Tolerances \texttt{rtol} $= 10^{-6}$ and \texttt{atol} $= 10^{-12}$ were used in stopping criterion \eqref{eq:stoppingcriterion}.  The initial iterate was set to $w=\max\{0,\psi\}$.

Table \ref{tab:results:classical} shows the resulting iteration counts for finest meshes of $m_J$ nodes.  When the RS Newton smoother is applied by itself as a solver (``RS only'') it shows rapidly-increasing iterations as expected, roughly proportional to resolution.  By contrast, V-cycle iterations increase slowly, perhaps logarithmically with the resolution.  Noting that for F-cycles we are counting the number of V-cycles after the ramp (Section \ref{sec:implementation}), we see that a constant number of cycles solves the problem.  Comparing the ball and spiral results, the complexity of the coincidence set or free boundary does not strongly-influence FASCD performance.  Note that each V-cycle does the work of about eight finest-level preconditioned Krylov (CG+IC) iterations.  Thus for the F-cycle results, where three iterations suffice at all resolutions, we see close to textbook multigrid efficiency\footnote{Textbook multigrid efficiency is imperfectly defined, but consider \cite{BrownSmithAhmadia2013}: ``an order of magnitude reduction in residual per iteration and solution of the fine-level system at a small multiple of the cost of a residual evaluation.''  Work of less than ten residual evaluations is often mentioned as the standard.} despite the low solution regularity.
\end{example}

%REGENERATE using git@bitbucket.org:pefarrell/fascd.git, in examples/
%$ timer python3 drape.py -prob ball|spiral -levs 9 -o ball|spiral.pvd -fascd_rtol 1.0e-6 -fascd_atol 1.0e-6 -fascd_cycle_type full
%$ paraview ball|spiral.pvd
%use preset "xray" and invert colors
\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.35\textwidth]{fixfigs/ball-set.png} \qquad \includegraphics[width=0.35\textwidth]{fixfigs/spiral-set.png}
\end{center}
\caption{Coincidence sets for the ball and spiral problems in Example \ref{ex:results:classical}.}
\label{fig:results:classical}
\end{figure}

@@

\begin{example}  \label{ex:results:plap}
Recall Example \ref{ex:plaplacian}, the nonlinear $p$-Laplacian operator.  Here we consider a unilateral obstacle problem, in one dimension for visualizability, for both fast ($p=1.5$) and degenerate ($p=4.0$) diffusivity cases.\footnote{``Fast'' and ``degenerate'' refer to $D=|\grad u|^{p-2}\to \infty$ and $D\to 0$ as $|\grad u|\to 0$, respectively; ellipticity is lost in the latter case.  Well-posedness and ($p$-dependent) solution regularity for $p$-Laplacian obstacle problems is considered in \cite{ChoeLewis1991}.}  This example reveals the sometimes strong contrast between $V(1,0)$ and $V(0,1)$ results (\pr{up}, \pr{down} $=0$ in Algorithm \ref{alg:fascd}, respectively), and it emphasizes the need for robust smoothers.

Suppose $\Omega=(-3,3)$, $\psi(x) = -0.2|x|$, $\mathcal{K} = \{v \ge \psi\} \subset W^{1,p}(\Omega)$, and $\ell(v) = \int_\Omega g v$ where $g(x)=1$ for $|x|<1$ and $g(x)=-1$ for $|x|>1$.  The exact solution of VI \eqref{eq:vi} is easily computed for any $p$; see Figure \ref{fig:results:plap}.  For this experiment we used a coarsest mesh of $m_0=7$ nodes and an $m_9=3073$ finest mesh.  The default smoother uses three (fixed) iterations of RS Newton, with LU solution of the step equations.\footnote{This is an optimal $O(m_j)$ smoother in 1D.}  The coarse solver is the same but iterated to convergence.  The initial iterate was $w=\psi$, and \texttt{rtol} $= 10^{-6}$ and \texttt{atol} $= 10^{-12}$.  Figure \ref{fig:imagesvcycle} illustrates the major steps of an example V-cycle, showing each of the coarse-correction VI subproblems and solutions.

@@

Solvers implemented in Firedrake can exploit parallel solvers from the PETSc library \cite{Balayetal2023}.  The above Examples, run in serial so as to fix the smoother,\footnote{The action of parallel preconditioners generally depends on the number of processes \cite[for example]{Bueler2021}.} also run efficiently in parallel.  The next Example applies a parallel solver, and Example \ref{ex:results:sia} demonstrates the parallel weak scaling \cite{Bueler2021} of FASCD.

>>>>

%\newcommand{\XX}{{\sffamily X}}
\newcommand{\XX}{\ding{55}}

@@

To allow box constraints on all meshes in the hierarchy, let $\tilde{\mathcal{V}}^j$ denote the set of functions on the vertices of $\mathcal{T}^j$ with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.  Suppose that the finest-level \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$ satisfy
\begin{equation}
\underline{\gamma}^J \le \overline{\gamma}^J, \qquad \underline{\gamma}^J < +\infty, \qquad -\infty < \overline{\gamma}^J. \label{eq:fe:boxconstraintrequirements}
\end{equation}
\pef{Shouldn't there be some relation between the continuous bounds and the discrete ones?}
In analogy to the continuous data, also assume that $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ on those vertices of $\mathcal{T}^J$ which lie in $\partial_D \Omega$.  Define
\begin{equation}
\mathcal{K}^J = \left\{w\,:\,\underline{\gamma}^J \le w \le \overline{\gamma}^J \text{ on vertices of } \mathcal{T}^J, \, \text{ and } \, w|_{\partial_D\Omega} = g_D^J|_{\partial_D\Omega}\right\} \subset \mathcal{V}^J, \label{eq:fe:fineconstraintset}
\end{equation}
the finest-level \emph{constraint set} of admissible FE functions. This is a nonempty, closed, and convex subset of $\mathcal{V}^J$.  %As $\mathcal{V}^J$ functions never take values $\pm\infty$, inequality in \eqref{eq:fe:fineconstraintset} is strict where $\underline{\gamma}^J, \overline{\gamma}^J$ are infinite; there is no lower or upper constraint, respectively, at such vertices.

@@

While \eqref{eq:fe:vi} approximates \eqref{eq:boxdirichletvi}, we warn the reader that it may not technically be conforming for that VI.  That is, even if quadrature and other implementation details for $f^J$ are ignored, and despite the vector space inclusion $\mathcal{V}^J \subset \mathcal{V}$, we do not require or generally expect $\mathcal{K}^J \subset \mathcal{K}$.  Regardless of the smoothness of the continuum obstacles $\underline{\gamma}, \overline{\gamma}$, admissibility with respect to their FE interpolants $\underline{\gamma}^J, \overline{\gamma}^J$ will generally not imply continuum admissibility.\footnote{This caveat is especially relevant in geophysical situations where only nodal values of the obstacle are known.  Nodal elevation data generally misses the actual topographic detail in-between the vertices; compare \cite{Bueler2016}.}

@@

To define injection $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$, observe that the vertices of $\mathcal{T}^{j-1}$ are vertices of $\mathcal{T}^j$, and that for any values at $\mathcal{T}^j$ vertices there is a unique $\mathcal{V}^j$ function assuming those values.  For $z\in\mathcal{V}^j$ we define the injection $\iR z$ as the element of $\mathcal{V}^{j-1}$ with the same (coarse) nodal values.  Injection is monotonic: if $w \le z$ then $\iR w \le \iR z$.

\pef{I don't like the name ``monotone restrictions''. The word restriction is strongly associated with operating on the dual space. I would prefer to call them ``monotone injections'', since that is the matching operator.}
The monotone restrictions $\maxR,\minR$ \cite{GraeserKornhuber2009}, which will only be applied to nodal functions $\tilde{\mathcal{V}}^j$, are nonlinear maps which maximize and minimize, respectively, nodal values over the open supports of coarser-mesh basis functions (i.e.~over the \emph{stars} of the coarse vertices). These operators are constructed to enforce the property
\begin{equation}
\minR z \le z \le \maxR z,  \label{eq:fe:monotonerestrictionprops}
\end{equation}
which does not hold for the standard injection operator $\iR$.
To make the construction precise, let $x_q^j$ denote the vertices of mesh $\mathcal{T}^j$, with corresponding basis functions $\psi_q^j(x)$.  For a $\tilde\RR$-valued nodal function $z\in\tilde{\mathcal{V}}^j$ we define functions in $\tilde{\mathcal{V}}^{j-1}$:
\begin{align}
(\maxR z)(x_p^{j-1}) &= \max \{z(x_q^j) \,:\, \psi_p^{j-1}(x_q^j) > 0\}, \label{eq:fe:monotonerestrictions} \\
(\minR z)(x_p^{j-1}) &= \min \{z(x_q^j) \,:\, \psi_p^{j-1}(x_q^j) > 0\}. \notag
\end{align}
For standard refinement a coarse-mesh node $x_p^{j-1}$ is in the set of fine vertices $x_q^j$ such that $\psi_p^{j-1}(x_q^j)>0$, so \eqref{eq:fe:monotonerestrictionprops} follows.
On the other hand, $z\ge 0$ implies $\minR z \ge 0$ and $z \le 0$ implies $\maxR z \le 0$.  These ordering properties are used in the CD construction of the next two Sections.
\pef{It would be good to have a diagram here. I could write one in tikz, if you don't have one handy.}

@@

Let $\Omega = (0,L) \subset \RR^1$ and consider a two-level mesh hierarchy ($J=1$).  Let $\underline{\gamma}^1$ be a sawtooth function with minimum value zero, and suppose $\overline{\gamma}^1=C+\underline{\gamma}^1$ with $C>0$ which is strictly less than the maximum of $\underline{\gamma}^1$.  Figure \ref{fig:directRbad} shows such functions; here levels $\mathcal{T}^1$ and $\mathcal{T}^0$ have $5$ and $3$ vertices, respectively.  Ignoring Dirichlet boundary for simplicity, the constraint set $\mathcal{K}^1$ defined by \eqref{eq:fe:fineconstraintset} is nonempty.  However, various restriction schemes applied directly to $\underline{\gamma}^1,\overline{\gamma}^1$ are problematic.  Suppose we propose
    $$\underline{\gamma}^0, \overline{\gamma}^0 \equiv \iR \underline{\gamma}^1, \iR \overline{\gamma}^1. \qquad \text{\emph{[problematic]}}$$
Then $\underline{\gamma}^0=0$ and $\overline{\gamma}^0=C$ identically.  In that case the corresponding constraint set $\mathcal{K}^0$ is nonempty, but prolongation $Pw^0$ of an admissible solution of the coarse problem ($\underline{\gamma}^0 \le w^0 \le \overline{\gamma}^0$) is not admissible on the finer level: $Pw^0 \notin \mathcal{K}^1$.  On the other hand, suppose we propose
    $$\underline{\gamma}^0, \overline{\gamma}^0 \equiv \maxR \underline{\gamma}^1, \minR \overline{\gamma}^1. \qquad \text{\emph{[problematic]}}$$
The general motivation for this choice---compare \eqref{eq:fe:chilevels} below---is that $w^0 \ge \underline{\gamma}^0 \, \implies \, Pw^0 \ge \underline{\gamma}^1$ and $w^0 \le \overline{\gamma}^0 \, \implies \, Pw^0 \le \overline{\gamma}^1$.  However, while these facts are true, the difficulty with this choice is that the coarse-level admissible set $\mathcal{K}^0$ is now empty because the proposed lower bound $\underline{\gamma}^0$ actually exceeds the proposed upper bound $\overline{\gamma}^0$.
\end{example}

@@

\begin{example}  \label{ex:notfullcd}
Let $a > 0$ be an arbitrary positive number.  Consider a 2-level hierarchy ($J=1$), and suppose the lower defect constraint is the constant function $\underline{\chi}^1=-a$.  Note $\underline{\chi}^0=\maxR \underline{\chi}^1=-a$ also, so $\underline{\phi}^1=0$ and $\underline{\phi}^0=-a$.  Suppose $z^1$ is a ``sawtooth'' function which takes on value $-a$ on the coarse vertices and value $0$ on the other vertices.  Then, for any $\overline{\chi}^1\ge 0$, we have $\overline{\chi}^1 \ge 0 \ge z^1\ge \underline{\chi}^1$, so $z^j \in \mathcal{U}^1$.  However, $I_{1\to 0}^\ominus(z^1 - \underline{\chi}^1) = \minR(z^1 - \underline{\chi}^1) = 0$ identically.  From \eqref{eq:fe:unilateraldecompositionoperator}, $\Pi_1 z^1$ simplifies to $\Pi_1 z^1 = z^1 - \underline{\chi}^1$.  But then $\Pi_1 z^1$ attains a maximum $a\ge 0$.  This maximum can be chosen to exceed any upper defect obstacle $\overline{\phi}^1$, so $\Pi_1 z^1 \notin \mathcal{D}^1$.
\end{example}

@@

Because the problem is inequality-constrained, repeated application of Algorithm \ref{alg:fascd} to solve the finite-dimensional, box-constrained VI \eqref{eq:fe:vi} will not generally reduce the residual $f^J(w^J) - \ell^J$ to zero everywhere.  To construct appropriate convergence criteria for the discretization, note \eqref{eq:fe:vi} is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003}.  Let $x_p \in \mathcal{T}^J$ be a finest-level node, and $\psi_p(x)$ the corresponding basis function.  Let $\mathcal{X}^J=\big\{w\in\cV^J\,:\,w(x_p)=g_D^J(x_p) \text{ if } x_p\in \partial_D\Omega\big\}$, thus $\cK^J = \left\{\underline{\gamma}^J \le w \le \overline{\gamma}^J\right\} \cap \mathcal{X}^J$.  If $m_D$ is the number of vertices $x_p$ in the Dirichlet boundary, let $I_D$ be the $m_D\times m_J$ matrix which extracts Dirichlet nodal values; its transpose $I_D^\top$ injects Dirichlet boundary degrees of freedom into $(\mathcal{V}^J)'$.  Finally, write $a\perp b$ when $a,b \in \RR^n$ are \emph{complementary}, that is, when $a_i \ge 0$ and $b_i \ge 0$ and $a_i b_i = 0$ for $i=0,\dots,n-1$.  Then problem \eqref{eq:fe:vi}, which would be expressed VI$\left(f^J(\cdot)-\ell^J,\left[\,\underline{\gamma}^J,\overline{\gamma}^J\right] \cap \mathcal{X}^J\right)$ in the operations research literature \cite{FerrisPang1997}, is equivalent to the following MCP for the solution $u^J \in \mathcal{K}^J$, but with additional equality constraints:
\begin{align}
\underline{\gamma}^J \le u^J \le \overline{\gamma}^J \quad &\perp \quad f^J(u^J) - \ell^J - I_D^\top \lambda \label{eq:fe:mcp} \\
I_D u^J &= g_D^J\big|_{\partial_D\Omega} \notag
\end{align}
for some Lagrange multipliers $\lambda \in \RR^{m_D}$.
\pef{I wonder if this definition will confuse people, because they won't know how to plug in $\underline{\gamma}^J \le u^J \le \overline{\gamma}^J$ into your definition of $\perp$ ($\underline{\gamma}^J \le u^J \le \overline{\gamma}^J$ is not in $\RR^n$). Maybe it would be easier to just use the spelled-out version below, and not use this notation?}

@@

In summary, $\rNC : \mathcal{K}^J \to \RR^{m_J}$ is a computable nonlinear map defined using the discretization $f^J$ and the vectors $\ell^J$, $\underline{\gamma}^J$, and $\overline{\gamma}^J$.  A value $\rNC(w^J)_p$ is nonzero only at vertices where \eqref{eq:fe:mcp} is violated, either because $r(w^J)_p$ remains nonzero where the constraints are inactive, or because it has the wrong sign at an active node, or because the boundary condition is not satisfied.

@@

Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} were implemented in Python using the Firedrake library \cite{Rathgeberetal2016}, allowing $P_1$ or $Q_1$ elements and based on standard nested mesh refinement.  Here we show numerical results for the example problems presented in Section \ref{sec:vi}, with a focus on the scaling of FASCD iterations as the mesh resolution increases.  Our main observations are that each V-cycle dramatically reduces the residual norm, and that very few F-cycles solve most examples to within discretization error.

\begin{example} \label{ex:results:classical}
We consider the unilateral 2D Laplacian obstacle problem presented in Example \ref{ex:plaplacian} with $p=2$ and admissible set $\mathcal{K} = \{v \ge \psi\} \subset W^{1,2}(\Omega)$.  We consider two problems with very different coincidence (active) sets $\{x\in\Omega \,:\, u(x)=\psi(x)\}$, both over square $\Omega$.   The standard ``ball'' problem has an upper hemisphere obstacle $\psi$ \cite[Chapter 12]{Bueler2021}.  In the ``spiral'' problem \cite[Example 7.1.1]{GraeserKornhuber2009}, $\psi$ is constructed to produce a more interesting coincidence set; see Figure \ref{fig:results:classical}.

For this experiment we used a coarsest mesh of $m_0=5^2$ vertices (32 triangles).  The finest mesh had $m_9=2049^2 \approx 4.2 \times 10^{6}$ vertices.  As our smoother we applied a single RS Newton iteration (Section \ref{sec:implementation}), with three (fixed) IC-preconditioned CG iterations to approximately solve the arising linear systems.  The coarse solver was RS$+$LU, iterated to convergence.  Tolerances \texttt{rtol} $= 10^{-6}$ and \texttt{atol} $= 10^{-12}$ were used in stopping criterion \eqref{eq:stoppingcriterion}.  The initial iterate was set to $w=\max\{0,\psi\}$.

Table \ref{tab:results:classical} shows the resulting iteration counts for finest meshes of $m_J$ vertices.  When the RS Newton smoother is applied by itself as a solver (``RS only'') it shows rapidly-increasing iterations as expected, roughly proportional to resolution.  By contrast, V-cycle iterations increase slowly, perhaps logarithmically with the resolution.  Noting that for F-cycles we are counting the number of V-cycles after the ramp (Section \ref{sec:implementation}), we see that a constant number of cycles solves the problem.  Comparing the ball and spiral results, the complexity of the coincidence set or free boundary does not appear to strongly influence FASCD performance \pef{unlike \cite{GraeserKornhuber2009}?}. Each V-cycle does the work of about eight finest-level preconditioned Krylov (CG+IC) iterations; for the F-cycle results, where three iterations suffice at all resolutions, we see close to textbook multigrid efficiency\footnote{Textbook multigrid efficiency is imperfectly defined, but consider \cite{BrownSmithAhmadia2013}: ``an order of magnitude reduction in residual per iteration and solution of the fine-level system at a small multiple of the cost of a residual evaluation.''  Work of less than ten residual evaluations is often mentioned as the standard.} despite the low solution regularity.
\end{example}

%REGENERATE using git@bitbucket.org:pefarrell/fascd.git, in examples/
%$ timer python3 drape.py -prob ball|spiral -levs 9 -o ball|spiral.pvd -fascd_rtol 1.0e-6 -fascd_atol 1.0e-6 -fascd_cycle_type full
%$ paraview ball|spiral.pvd
%use preset "xray" and invert colors
\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.35\textwidth]{fixfigs/ball-set.png} \qquad \includegraphics[width=0.35\textwidth]{fixfigs/spiral-set.png}
\end{center}
\caption{Coincidence sets for the ball and spiral problems in Example \ref{ex:results:classical}.}
\label{fig:results:classical}
\end{figure}
\pef{The spiral problem looks under-resolved; is that right?}

@@

\begin{example}  \label{ex:results:plap}
We again consider Example \ref{ex:plaplacian} with a unilateral constraint, but now for the nonlinear $p$-Laplacian. We investigate both fast ($p=1.5$) and degenerate ($p=4.0$) diffusivities.\footnote{``Fast'' and ``degenerate'' refer to $|\grad u|^{p-2}\to \infty$ and $\to 0$, respectively; ellipticity is lost in the latter case.  Well-posedness and ($p$-dependent) solution regularity for $p$-Laplacian obstacle problems is considered in \cite{ChoeLewis1991}.} We consider the problem in one dimension for ease of visualization. This example reveals the sometimes strong contrast between $V(1,0)$ and $V(0,1)$ results (\pr{up}, \pr{down} $=0$ in Algorithm \ref{alg:fascd}, respectively), and it emphasizes the need for robust smoothers.

Let $\Omega=(-3,3)$, $\psi(x) = -0.2|x|$, $\mathcal{K} = \{v \ge \psi\} \subset W^{1,p}(\Omega)$, and $\ell(v) = \int_\Omega g v$ where $g(x)=1$ for $|x|<1$ and $g(x)=-1$ for $|x|>1$.  The exact solution of VI \eqref{eq:vi} is easily computed for any $p$; see Figure \ref{fig:results:plap}. \pef{Write down the exact solution?}  For this experiment we used a coarsest mesh of $m_0=7$ vertices and a finest mesh with $m_9=3073$.  The default smoother uses three (fixed) iterations of RS Newton, with LU solution of the arising linear systems.\footnote{This is an optimal $O(m_j)$ smoother in 1D.}  The coarse solver is the same but iterated to convergence.  The initial iterate was $w=\psi$, and \texttt{rtol} $= 10^{-6}$ and \texttt{atol} $= 10^{-12}$.  Figure \ref{fig:imagesvcycle} illustrates the major steps of an example V-cycle, showing each of the coarse-correction VI subproblems and solutions.

@@

Solvers implemented in Firedrake can exploit parallel solvers from the PETSc library \cite{Balayetal2023}.  The above Examples, run in serial so as to fix the smoother,\footnote{The action of parallel preconditioners can depend on the number of processes \cite[for example]{Bueler2021}.} also run efficiently in parallel.  The next Example applies a parallel solver, and Example \ref{ex:results:sia} demonstrates the parallel weak scaling \cite{Bueler2021} of FASCD.
